---
title: ""
author: "Madison Carpenter"
date: "2023-02-16"
output:
  html_document: default
  pdf_document: default
---

# 1.0 Setup the Environment

```{r Setup the Environment, echo=FALSE, message=FALSE}
# Load packages
library(tidyverse)
library(magrittr)
library(WriteXLS)
library(pathview)
library(xlsx)
library(ggVennDiagram)
library(factoextra)
library(gt)
library(clusterProfiler)
library(KEGGREST)

# Setup Pathways
workdir <- "XXXXXX"
scriptDir <- "XXXXX"

# Set version control
set.seed(123)

# # Load DESeq workflow scripts
source(file = paste0(workdir,scriptDir,"DESeq.R"))
source(file = paste0(workdir,scriptDir,"DataVisualizations.R"))
source(file = paste0(workdir,scriptDir,"Deseq_Statistics.R"))
source(file = paste0(workdir,scriptDir, "pathway_graphs.R"))
source(file = paste0(workdir,scriptDir, "GoEnrichAnalysis.R"))

# Set Diagram settings
options(width=200)
```

# 2.0 Load Data

```{r pressure, echo=FALSE}
#_____________UPDATE!______________#
# Setup Variables
expdir <- "XXXXXX"

meta <- readxl::read_excel(paste0(workdir,expdir,"Metadata.xlsx"))
quantfiles <- file.path(workdir,expdir,"data/", meta$Well, "quant.sf")
#_________________________________#
```

# 3.0 Figure 6C, and 6G
## 3.1 DESEQ
### 3.1.1 Modify Metadata

```{r}
# Extract Xenon Samples Meta Data
sub_meta <- meta %>%
  filter(Disease == "WT_PHAG" | Disease == "IFNcKO_PHAG") %>%
  dplyr::select(-Age) %>%
  separate(., col = "Disease",sep = "_", into = c("Genotype","PhagStatus")) %>% 
  mutate(Treatment = ifelse(Treatment == "XEN","Xenon",Treatment))
```

```{r}
##________UPDATE ENTIRE SECTION___________##
# Set a file prefix
fileprefix <- "Xenon_phago_IFNcKoXenonvsWTXenon" # Experimental group should be listed first in the file prefix**

# Extract Xenon Samples Meta Data
sub_meta2 <- sub_meta %>% filter(Treatment == "Xenon") %>% mutate(Comparision = paste0("IFNcKoXenonvsWTXenon"))

# Pull specific Quant files
position_list2 <- paste0(sub_meta2$Well, collapse = '|')
sample_files2 <- grep(position_list2, quantfiles, value = TRUE) # Check number of files matches # of samples

## Create Graphs All 
table_lab <- paste0(sub_meta2$Well,"_",sub_meta2$Gender,"_",sub_meta2$Treatment,"_",sub_meta2$Genotype)
```

### 3.1.2 Tximport/DESEQ

```{r, echo=FALSE}
#_____________UPDATE!______________#
# Setup the Design for Tximport/DESeq2
design = "Genotype"
treatcolumn = "Genotype"
control = "WT"
refanimal = "mouse"
type = "LRT"
#_________________________________#

# Run Tx2gene
tx2gene <- tx2genefun(reference = refanimal)

# Run Tximport and DESeq 
ddsTxi <- Tximportfun(sample_files2,reference = refanimal, metadata = sub_meta2, design = design, tx2gene = tx2gene)
dds <- DESeqFun(treatcolumn = treatcolumn, control = control, DESeqObject = ddsTxi,test_type = type)

# Save Deseq output
ddsobj <- dds[[1]] # DESeq Object
ddsres <- dds[[2]]# Results
```

### 3.1.3 Export Data

```{r, echo = FALSE}
#_____________UPDATE!______________#
resdir <- paste0(workdir,expdir,"XXXX")
#__________________________________#

# Extract the stats
extractedstats <- extractstats(ddsobj, ContrastObj = ddsres ,sub_meta2, graphlabel = table_lab)

# Save to variables
DEResults <- extractedstats[[1]] 
DEcounts <- extractedstats[[2]] 
sigDEResults <- extractedstats[[3]]
sigDEcounts <- extractedstats[[4]] 

# Export the Stats
exportstats(DEResults = DEResults, DEcounts = DEcounts,sigDEResults = sigDEResults, 
            sigDEcounts = sigDEcounts, directory = resdir,fileprefix = fileprefix)
```

## 3.2 Data Exploration
### ScatterPlot Volcano (Figure 6C)

```{r}
# Setup Volcano Scatter Theme
theme_scatter <- theme(aspect.ratio = 1, 
                       panel.background = element_blank(),
                       panel.border=element_rect(fill=NA),
                       panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank(),
                       strip.background=element_blank(),
                       axis.title  = element_text(size = 20, colour = "black"),
                       axis.text.x=element_text(colour="black",size =20),
                       axis.text.y=element_text(colour="black",size =20),
                       axis.ticks=element_line(colour="black"),
                       plot.margin=unit(c(1,1,1,1),"line"),
                       plot.title = element_text(colour="black",size =20))
# Setup Variables 
volcano <- DEResults %>% .[-1,]
volcano$log_pval <- -log10(volcano$pvalue)
sig_up_data <- volcano %>% filter(log2FoldChange > 0) %>% filter(pvalue < 0.05)
sig_down_data <- volcano %>% filter(log2FoldChange < 0) %>% filter(pvalue < 0.05)

# Setup Genes of Interest
gene.of.interest <- c("Ccl6","Ifngr1","Cxcl9","Clec12a","Jund","Klf4","Ing2","Ccl24","Nlrp3","Plau","Hif3a","Ms4a1","Lif","Apob","Fgf2","Mrc2")
UP_genes_data <- sig_up_data %>% filter(gene %in% gene.of.interest)
DOWN_genes_data <- sig_down_data %>% filter(gene %in% gene.of.interest)



# Scatter Volcano Plot
scatter <- ggplot(volcano, aes(x= log2FoldChange, y= log_pval, label = gene)) +
  geom_point(size=0.5, color = 'grey90') +
  geom_point(data = sig_up_data, color = 'red', size = 1)+
  geom_point(data = sig_down_data, color = 'dodgerblue3', size = 1) +
  theme_scatter +
  xlim(-5,5)+
  ylim(0,25) +
  geom_vline(xintercept = 0.15,colour="grey", linetype = "longdash") +
  geom_vline(xintercept = -0.15,colour="grey", linetype = "longdash") +
  geom_hline(yintercept = 1.3, colour="grey", linetype = "longdash") +
  labs(title = table_header,
       x = "Log2FC",
       y = "-log(pvalue)") +
    geom_text_repel(data = UP_genes_data,
                  color = 'black', size = 6,fontface = 'italic',
                  min.segment.length = 0,
                  nudge_y = 5,
                  nudge_x = 4 - abs(UP_genes_data$log2FoldChange),
                  max.overlaps = Inf,
                  box.padding = unit(0.5, 'mm'),
                  point.padding = unit(0.5, 'mm'),
                  direction = "y") +
  geom_text_repel(data = DOWN_genes_data,
                   color = 'black',size = 6, fontface = 'italic',
                   min.segment.length =0,
                   nudge_y = 3,
                   nudge_x = -4 + abs(DOWN_genes_data$log2FoldChange),
                   max.overlaps = Inf,
                   point.padding = unit(0.5, 'mm'),
                  direction = "y")

# Save Figure
ggsave(filename = paste0(workdir, expdir, "plots/VolcanoPlots/", fileprefix, "GOI.pdf"),
  plot = scatter,
  device = NULL,
  path = NULL,
  scale = 1,
  width = NA,
  height = NA,
  units = c("mm"),
  dpi = 300,
  limitsize = TRUE,
  bg = NULL
)
```

### Generate IPA DotPlot (Figure 6G)
```{r}
##________UPDATE ENTIRE SECTION___________##
IPAgraphdir <- paste0(workdir,expdir,"plots/IPA/Phagocytic/")
##________________________________________##

# Load Data of Interst
IFNcKoXenonvsWTXenon <- readxl::read_excel(paste0(workdir,expdir,"results/Madison/Phagocytic/Xenon_phago_IFNcKoXenonvsWTXenon_statistics_pval01_redo_pathways.xls"), skip = 1) %>% 
  dplyr::mutate(Experiment = paste0("IFNcKoXenonvsWTXenon")) %>% 
  dplyr::rename(c("Ingenuity.Canonical.Pathways" = "Ingenuity Canonical Pathways", "z.score" = "z-score"))

# Setup Pathways of interest
path.of.interest <- c(
"Oxidative Phosphorylation",
"EIF2 Signaling",
"RHOA Signaling",
"LXR/RXR Activation",
"IL-15 Signaling",
"NF-κB Signaling",
"Phagosome Formation",
"ERK/MAPK Signaling",
"Pyroptosis Signaling Pathway",
"S100 Family Signaling Pathway",
"Mitochondrial Dysfunction")

# Generate Graph of Interest
IPApoigraph(data = IFNcKoXenonvsWTXenon, POI = path.of.interest, directory = IPAgraphdir, fileprefix = fileprefix, width = 6, height =8)
```


# Figure 6: 
## 5.1 DESEQ
### 5.1.1 Modify Data

```{r}
#### Extract Xenon Samples Meta Data
fileprefix <- "Xenon_phago_AllComparisions"
sub_meta <- meta %>%
  filter(Disease == "WT_PHAG" | Disease == "IFNcKO_PHAG") %>%
  dplyr::select(-Age) %>%
  separate(., col = "Disease",sep = "_", into = c("Genotype","PhagStatus")) %>% 
  mutate(Treatment = ifelse(Treatment == "XEN","Xenon",Treatment)) %>%
  mutate(Group = paste0(Genotype,"_",Treatment))

# Pull specific Quant files
position_list <- paste0(sub_meta$Well, collapse = '|')
sample_files <- grep(position_list, quantfiles, value = TRUE) # Check number of files matches # of samples

## Create Graphs All 
table_header <- "All Comparisions"
graph_lab <- paste0(sub_meta$Well,"\n",sub_meta$Gender,"\n",sub_meta$Treatment,"\n", sub_meta$Genotype) # Assign Rowname label
table_lab <- paste0(sub_meta$Well,"_",sub_meta$Gender,"_",sub_meta$Treatment,"_",sub_meta$Genotype)
```

### 5.1.2 TXIMPORT/DESEQ

```{r, echo=FALSE}
#_____________UPDATE!______________#
# Setup the Design for Tximport/DESeq2
design = "Group"
treatcolumn = "Group"
control = "WT_AIR"
refanimal = "mouse"
var1 = "IFNcKO_Xenon"
var2 = "WT_AIR"
type = "LRT"
#_________________________________#

# Run Tx2gene
tx2gene <- tx2genefun(reference = refanimal)

# Run Tximport and DESeq 
ddsTxi1 <- Tximportfun(sample_files,reference = refanimal, metadata = sub_meta, design = design, tx2gene = tx2gene)
dds1 <- DESeqFun(treatcolumn = treatcolumn, control = control, DESeqObject = ddsTxi1, var1 = var1, var2 = var2, test_type = type)

# Save Deseq output
ddsobj <- dds1[[1]] # DESeq Object
ddsres <- dds1[[2]]# Results
```

### 5.1.3 Export Stats

```{r, echo = FALSE}
#_____________UPDATE!______________#
resdir <- paste0(workdir,expdir,"XXXX")
#__________________________________#

# Extract the stats
extractedstats <- extractstats(ddsobj, ContrastObj = ddsres ,sub_meta, graphlabel = table_lab)

# Save to variables
DEResults <- extractedstats[[1]] 
DEcounts <- extractedstats[[2]] 
sigDEResults <- extractedstats[[3]]
sigDEcounts <- extractedstats[[4]] 

# Export the Stats
exportstats(DEResults = DEResults, DEcounts = DEcounts,sigDEResults = sigDEResults, 
            sigDEcounts = sigDEcounts, directory = resdir,fileprefix = fileprefix)
```

## 5.2 Data Exploration
### 5.2.1 Heatmap (Figure G)

```{r, echo=FALSE, fig.show='hide'}
##________Update_______##
heatmapdir <- paste0(workdir,expdir,"XXXX")
#______________________#

# Generate Heatmap for significant genes
map <- genheatmap(sigDEcounts, colnames(sigDEcounts)[-1], height = 310, width = 15, directory = heatmapdir, cutrow = 4, gap_col = c(4,8,12),fileprefix = fileprefix, suffix = "pvalue05.pdf", show_row = TRUE)

genheatmap(sigDEcounts, colnames(sigDEcounts)[-1], height = 310, width = 15, directory = heatmapdir, cutrow = 4, gap_col = c(4,8,12),fileprefix = fileprefix, suffix = "pvalue05_nolabels.pdf", show_row = FALSE)
```

## 5.3 GSEA Analysis
### 5.3.1 Step 1: Extract Clusters
```{r}
##________Update_______##
heatmap <- map
reforg <- "mouse"
numclusters <- "4"
GSEAdir <- paste0(workdir,expdir,"plots/GSEAGO/")
#______________________#

# Extract Clusters from Dendrogram
clusters <- ExtClustersfun(heatmap,numclusters = numclusters)

# Export Cluster ID per gene 
xlsx::write.xlsx(x = clusters, file = paste0(workdir, expdir, "results/", fileprefix, "ClusterID_GSEAGO_4breaks.xlsx"), row.names = FALSE)
```

### 5.3.2 Step2: GO Enrichment Analysis

```{r}
# Run GO Enrichment Analysis
GSEAGOresults <- GSEAGOAnalysis(sigDEResults,geneclusters = clusters ,reforg = reforg)

# GSEA go results per cluster
Clust1obj <- GSEAGOresults[[1]]
Clust2obj <- GSEAGOresults[[2]]
Clust3obj <- GSEAGOresults[[3]]
Clust4obj <- GSEAGOresults[[4]]

# Save results
Clust1res <- Clust1obj %>% .@result %>% filter(pvalue < 0.1)
Clust2res <- Clust2obj %>% .@result %>% filter(pvalue < 0.1)
Clust3res <- Clust3obj %>% .@result %>% filter(pvalue < 0.1)
Clust4res <- Clust4obj %>% .@result %>% filter(pvalue < 0.1)

# Export GSEAGO results
xlsx::write.xlsx(x = Clust1res, file = paste0(workdir, expdir, "results/", fileprefix, "Cluster1_GSEAGOres.xlsx"), row.names = TRUE)
xlsx::write.xlsx(x = Clust2res,file = paste0(workdir, expdir, "results/", fileprefix, "Cluster2_GSEAGOres.xlsx"), row.names = TRUE)
xlsx::write.xlsx(x = Clust3res, file = paste0(workdir, expdir, "results/", fileprefix, "Cluster3_GSEAGOres.xlsx"), row.names = TRUE)
xlsx::write.xlsx(x = Clust4res,file = paste0(workdir, expdir, "results/", fileprefix, "Cluster4_GSEAGOres.xlsx"), row.names = TRUE)
```

### 5.3.3 Step3: Generate Enrichment Graphs (Figure H-I)

```{r}
# Set the GSEA directory 
GSEAdir <- paste0(workdir,expdir,"plots/GSEAGO/")


# Setup Pathways of Interst
poi_cluster1 <- c("cellular response to tumor necrosis factor",
"lipid homeostasis",
"regulation of mitochondrion organization",
"regulation of myeloid leukocyte differentiation",
"DNA biosynthetic process",
"nervous system process",
"cell cycle",
"cell surface receptor signaling pathway")


poi_cluster2 <- c("glial cell differentiation",
"cell surface receptor signaling pathway involved in cell-cell signaling",
"intracellular protein transport",
"mRNA destabilization",
"negative regulation of cell migration",
"cytokine-mediated signaling pathway",
"positive regulation of NF-kappaB transcription factor activity",
"negative regulation of cell motility")


# Generate Pathway Graphs
Enrichgenes(Clust1obj,table_header = table_header,clustnum = "1",directory = GSEAdir,fileprefix = fileprefix, numshown = poi_cluster1)
Enrichgenes(Clust2obj,table_header = table_header,clustnum = "2",directory = GSEAdir,fileprefix = fileprefix, numshown = poi_cluster2)

# Export ClusterID data
sigDEResults_mod <- merge(sigDEResults, clusters, by = "gene")
xlsx::write.xlsx(x = sigDEResults_mod, file = paste0(workdir, expdir, "results/", fileprefix, "_sigResults_GSEAGO_ClusterID.xlsx"), row.names = FALSE)
```

