---
title: "SK5A4G_DESEQ"
author: "Madison Carpenter"
date: "2023-06-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Dropbox (Partners HealthCare)/")
setwd("~/Dropbox (Partners HealthCare)")
```

# 1.0 Setup the Environment

```{r Setup the Environment, echo=FALSE, message=FALSE}

suppressPackageStartupMessages({
# Load packages
library(tidyverse)
library(magrittr)
library(WriteXLS)
library(xlsx)
library(ggVennDiagram)
library(factoextra)
library(gt)
library(org.Mm.eg.db)
library(clusterProfiler)
library(ggpubr)
})

# Setup Pathways
workdir <- "~/Dropbox (Partners HealthCare)/"
scriptDir <- "Scripts/BulkRNAseq/"

# Set version control
set.seed(123)

# # Load DESeq workflow scripts
source(file = paste0(workdir,scriptDir,"DESeq.R"))
source(file = paste0(workdir,scriptDir,"DataVisualizations.R"))
source(file = paste0(workdir,scriptDir,"Deseq_Statistics.R"))
source(file = paste0(workdir,scriptDir, "pathway_graphs.R"))
source(file = paste0(workdir,scriptDir, "GoEnrichAnalysis.R"))

# Set Diagram settings
options(width=200)
```

# 2.0 Figure 3M, P and R
## 2.1 Load Data
```{r pressure, echo=FALSE}
#_____________UPDATE!______________#
# Setup Variables
expdir <- "SK-5A4G_SN0267744_Wesley_Nov_2022_Xenon/"

# Load Data
meta <- readxl::read_excel(paste0(workdir, expdir,"metadata_SK-5A4G.xlsx")) %>% dplyr::rename("Position" = "Well")
quantfiles <- file.path(workdir,expdir,"data/", meta$Position, "quant.sf")
#_________________________________#
```

## 2.2 DESEQ
### 2.2.1 Modify Metadata 
```{r, echo=FALSE}
##________UPDATE ENTIRE SECTION___________##
# Set a file prefix
fileprefix <- "SK-5A4G_H5XFAD_6M_Astrocytes" # Experimental group should be listed first in the file prefix**

# Extract Samples Meta Data & Filter Outliers
sub_meta <- meta %>% filter(Age == "6M" & Celltype == "AST") %>% mutate(Grouped = paste0(Position,"_",Treatment, "_", Gender))

# Pull specific Quant files
position_list <- paste0(sub_meta$Position, collapse = '|')
sample_files <- grep(position_list, quantfiles, value = TRUE) # Check number of files matches # of samples

# Check the metadata and files are in the correct order
## QC check - Order of metafile samples must match quant file order!  Fix metadata file if not!
all(file.exists(sample_files))

## Create Graphs All 
graph_lab <- sub_meta$Grouped

# Format the Meta Data
table_header <- "Human 5XFAD 6M Astocytes"
sub_meta %>%
  gt() %>% 
  tab_header(
    title = md("**Meta Data for Female Monocytes SK-5A4G samples**"),
    subtitle = md(paste0(table_header, " experiment's meta data are presented"))
  ) %>%
  cols_move_to_start(
    columns = c(Position, Treatment) # Update based on columns **
  )
```

### 2.2.2 Tximport/DESeq

```{r, echo=FALSE}
#_____________UPDATE!______________#
# Setup the Design for Tximport/DESeq2
design = "Treatment"
treatcolumn = "Treatment"
control = "AIR"
refanimal = "mouse"
type <- "LRT"
#_________________________________#
# Run Tx2gene
tx2gene <- tx2genefun(reference = refanimal)

# Run Tximport and DESeq 
ddsTxi <- Tximportfun(sample_files,reference = refanimal, metadata = sub_meta, design = design, tx2gene = tx2gene)
dds <- DESeqFun(treatcolumn = treatcolumn, control = control, DESeqObject = ddsTxi, test_type = type)

# Save Deseq output
ddsobj <- dds[[1]] # DESeq Object
ddsres <- dds[[2]]# Results
```

### 2.2.3 Export DESEQ results
```{r}
#_____________UPDATE!______________#
resdir <- paste0(workdir,expdir,"results/Madison/Astrocytes_6months/")
#__________________________________#

# Extract the stats on the DESeq run 
extractedstats <-  extractstats(ddsobj, ContrastObj = ddsres ,sub_meta, graphlabel = graph_lab)

# Save to variables
DEResults <- extractedstats[[1]] 
DEcounts <- extractedstats[[2]] 
sigDEResults <- extractedstats[[3]]
sigDEcounts <- extractedstats[[4]] 

# Get Summary on DESEQ run
summary(ddsres, alpha = 0.05)

# Export the Stats
exportstats(DEResults = DEResults, DEcounts = DEcounts,sigDEResults = sigDEResults, 
            sigDEcounts = sigDEcounts, directory = resdir,fileprefix = fileprefix)
```

## 2.3 Data Exploration
### 2.3.1 Heatmap (Figure 3M)

```{r, echo=FALSE, fig.show='hide'}
##________Update_______##
heatmapdir <- paste0(workdir,expdir,"plots/Heatmaps/Astrocytes/")
#______________________#

# Generate HMAP for all significant genes
genheatmap(sigDEcounts, graph_lab = graph_lab, 
           height = 270, width = 10, suffix = paste0("_0.05pvalue.pdf"),
           directory = heatmapdir, fileprefix = fileprefix,
           cutrow = 2, gap_col = 5, show_row = TRUE)

# Generate HMAP for all significant genes - NO Label
genheatmap(sigDEcounts, graph_lab = graph_lab, 
           height = 270, width = 10, suffix = paste0("_NOLabel_0.05pvalue.pdf"),
           directory = heatmapdir, fileprefix = fileprefix,
           cutrow = 2, gap_col = 5, show_row = FALSE)
```

### 2.3.2 Volcano Plot (Figure 3P)

```{r}
# Setup Volcano Scatter Theme
theme_scatter <- theme(aspect.ratio = 1, 
                       panel.background = element_blank(),
                       panel.border=element_rect(fill=NA),
                       panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank(),
                       strip.background=element_blank(),
                       axis.title  = element_text(size = 20, colour = "black"),
                       axis.text.x=element_text(colour="black",size =20),
                       axis.text.y=element_text(colour="black",size =20),
                       axis.ticks=element_line(colour="black"),
                       plot.margin=unit(c(1,1,1,1),"line"),
                       plot.title = element_text(colour="black",size =20))
# Setup Variables 
volcano <- DEResults %>% .[-1,]
volcano$log_pval <- -log10(volcano$pvalue)
sig_up_data <- volcano %>% filter(log2FoldChange > 0) %>% filter(pvalue < 0.05)
sig_down_data <- volcano %>% filter(log2FoldChange < 0) %>% filter(pvalue < 0.05)

# Setup Variables for genes to label
DOWN_genes <- c("App","Crebbp","Grin2c","Inpp5d","Sod3")
DOWN_genes_data <- sig_down_data %>% filter(gene %in% DOWN_genes)

UP_genes <- c("Gfap","Serpina3f","Stat1")
UP_genes_data <- sig_up_data %>% filter(gene %in% UP_genes)

one_gene <- "Nr3c2"
one_gene_data <- sig_up_data %>% filter(gene %in% one_gene)


# Scatter Volcano Plot
scatter <- ggplot(volcano, aes(x= log2FoldChange, y= log_pval, label = gene)) +
  geom_point(size=0.5, color = 'grey90') +
  geom_point(data = sig_up_data, color = 'red', size = 1)+
  geom_point(data = sig_down_data, color = 'dodgerblue3', size = 1) +
  theme_scatter +
  xlim(-5,5)+
  #ylim(0,8) +
  geom_vline(xintercept = 0.15,colour="grey", linetype = "longdash") +
  geom_vline(xintercept = -0.15,colour="grey", linetype = "longdash") +
  geom_hline(yintercept = 1.3, colour="grey", linetype = "longdash") +
  labs(title = table_header,
       x = "Log2FC",
       y = "-log(pvalue)") +
    geom_text_repel(data = UP_genes_data,
                  color = 'black', size = 6,fontface = 'italic',
                  min.segment.length = 0,
                  nudge_y = 1.5,
                  nudge_x = 2 - abs(UP_genes_data$log2FoldChange),
                  max.overlaps = Inf,
                  box.padding = unit(0.5, 'mm'),
                  point.padding = unit(0.5, 'mm'),
                  direction = "y") +
  geom_text_repel(data = DOWN_genes_data,
                   color = 'black',size = 6, fontface = 'italic',
                   min.segment.length =0,
                   nudge_y = 2,
                   nudge_x = -3 + abs(DOWN_genes_data$log2FoldChange),
                   max.overlaps = Inf,
                   point.padding = unit(0.5, 'mm'),
                  direction = "y") +
  geom_text_repel(data = one_gene_data,
                   color = 'black',size = 6, fontface = 'italic',
                   min.segment.length =0,
                   nudge_y = 2,
                   nudge_x = 2 - abs(one_gene_data$log2FoldChange),
                   max.overlaps = Inf,
                   point.padding = unit(0.5, 'mm'),
                  direction = "y")

ggsave(filename = paste0(workdir, expdir, "plots/VolcanoPlots/", fileprefix, "GOI_redo2.pdf"),
  plot = scatter,
  device = NULL,
  path = NULL,
  scale = 1,
  width = NA,
  height = NA,
  units = c("mm"),
  dpi = 300,
  limitsize = TRUE,
  bg = NULL
)
scatter 
```

## 2.4 NicheNet Analysis
### 2.4.1 Load Microglia Data

```{r pressure, echo=FALSE}
# Load Data - Code to generate this data is listed in File: 
HM_sigDEResults <- xlsx::read.xlsx("~/Dropbox (Partners HealthCare)/SK-5A4F_SN0268732_Wesley_Nov_2022_Xenon_Human_Mouse/results/Madison/IPSMG_6M/SK-5A4F_IPMG_6M_statistics_pval05.xlsx", sheetName = "sigDEResults")
HM_sigDEcounts <- xlsx::read.xlsx("~/Dropbox (Partners HealthCare)/SK-5A4F_SN0268732_Wesley_Nov_2022_Xenon_Human_Mouse/results/Madison/IPSMG_6M/SK-5A4F_IPMG_6M_counts_pval05.xlsx", sheetName = "sigDEcounts")

# Assign Astrocytes counts and Results to new variables
astrsigDEResults <- sigDEResults
astro_sigDEcounts <- sigDEcounts
```

### 2.4.2 Load Ligand-Receptor Interaction Database

```{r}
# Import in the Ligand Information
ligand_target_matrix = readRDS(url("https://zenodo.org/record/3260758/files/ligand_target_matrix.rds"))
lr_network = readRDS(url("https://zenodo.org/record/3260758/files/lr_network.rds"))
weighted_networks = readRDS(url("https://zenodo.org/record/3260758/files/weighted_networks.rds"))
weighted_networks_lr = weighted_networks$lr_sig %>% inner_join(lr_network %>% distinct(from,to), by = c("from","to"))

# Convert Human to Mouse Network
lr_network = lr_network %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
colnames(ligand_target_matrix) = ligand_target_matrix %>% colnames() %>% convert_human_to_mouse_symbols()
rownames(ligand_target_matrix) = ligand_target_matrix %>% rownames() %>% convert_human_to_mouse_symbols()
ligand_target_matrix = ligand_target_matrix %>% .[!is.na(rownames(ligand_target_matrix)), !is.na(colnames(ligand_target_matrix))]

#Convert Weighted Network
weighted_networks_lr = weighted_networks_lr %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
```



```{r}
# Isolate out Gene of Interest per Cell Type 
HMicroglia_goi <- HM_sigDEcounts %>% mutate(gene = tools::toTitleCase(tolower(gene))) %>% filter(gene %in% rownames(ligand_target_matrix)) %>% pull(gene) %>% gsub("Lgals3bp", "Lgals3", .)
Astrocytes_goi <- astro_sigDEcounts$gene %>% .[. %in% rownames(ligand_target_matrix)] 
```


### 2.4.3 Step 1: Define set of potential ligands

```{r}
# Pull Receptor and Ligand information. Match with GOI of interest
ligands = lr_network %>% pull(from) %>% unique()
expressed_ligands = intersect(ligands,HMicroglia_goi)

receptors = lr_network %>% pull(to) %>% unique()
expressed_receptors = intersect(receptors,Astrocytes_goi)

lr_network_expressed = lr_network %>% filter(from %in% expressed_ligands & to %in% expressed_receptors)
potential_ligands = lr_network_expressed %>% pull(from) %>% unique()
```

### 2.4.4 Step 2: Perform NicheNet
```{r}
# Get Background Genes 
background_expressed_genes = Astrocytes_goi %>% .[. %in% rownames(ligand_target_matrix)]
# Predict Ligand Activity
ligand_activities = predict_ligand_activities(geneset = HMicroglia_goi, 
                                              background_expressed_genes = background_expressed_genes, ligand_target_matrix = ligand_target_matrix,
                                              potential_ligands = potential_ligands)

# Identify Best upstream ligands
best_upstream_ligands = ligand_activities %>% top_n(20, pearson) %>% arrange(-pearson) %>% pull(test_ligand)
```

### 2.4.5 Step 3: Infer target genes of top ranked ligands

```{r}
# Identify Active Ligand Links
active_ligand_target_links_df = best_upstream_ligands %>% 
  lapply(get_weighted_ligand_target_links, geneset = HMicroglia_goi, ligand_target_matrix = ligand_target_matrix, n = 250) %>% 
  bind_rows()

# Prepare visual
active_ligand_target_links = prepare_ligand_target_visualization(ligand_target_df = active_ligand_target_links_df, ligand_target_matrix = ligand_target_matrix, cutoff = 0.25)
```


```{r}
# get the ligand-receptor network of the top-ranked ligands
lr_network_top = lr_network %>% filter(from %in% best_upstream_ligands & to %in% expressed_receptors) %>% distinct(from,to)
best_upstream_receptors = lr_network_top %>% pull(to) %>% unique()

# get the weights of the ligand-receptor interactions as used in the NicheNet model
lr_network_top_df = weighted_networks_lr %>% filter(from %in% lr_network_top$from & to %in% lr_network_top$to)

# Get FoldChange information on Ligands
HMvsAnetwork <- lr_network_top_df
HM_sigDEResults$gene <- gsub("LGALS3BP", "Lgals3", HM_sigDEResults$gene)
ligandFC <- HM_sigDEResults %>% filter(tolower(gene) %in% tolower(HMvsAnetwork$from)) %>% select(gene,log2FoldChange) %>% rename("from" = "gene") %>% mutate(from = tools::toTitleCase(tolower(from)))
receptorFC <- astrsigDEResults %>% mutate(to = toupper(gene)) %>%filter(tolower(to) %in% tolower(HMvsAnetwork$to)) %>% select(c(to,log2FoldChange)) %>% mutate(to = tools::toTitleCase(tolower(to)))

# Merge FoldChange
HMvsAnetwork <- merge(x = lr_network_top_df, y = ligandFC, by.x = "from") %>% rename(., "L_L2FC" = "log2FoldChange")
HMvsAnetwork <-  merge(x = HMvsAnetwork, y = receptorFC, by.x = "to") %>% rename(., "R_L2FC" = "log2FoldChange")
HMvsAnetwork <- HMvsAnetwork[,c(2,1,3,4,5)] %>% rename("value" = "weight")
```

### 2.4.6 Step 4: Generate Circos Plot (Figure 3R)

```{r fig.width=5}
## Create a circos plot
x <- HMvsAnetwork %>% mutate(from = paste0("HM", from), to = paste0("AS",to))
sectors <- sort(unique(c(x$from, x$to)))
df.groups <- structure(substr(sectors, start = 1, stop = 2), names = sectors)
Type_Cols <- c(HM = "blue", AS = "red")

# Create Groups for Fold Change: Setting Colour scheme
xx <- x %>% mutate(from = paste0(ifelse(L_L2FC < -1, "Low2",
                                                   ifelse(L_L2FC > -1 & L_L2FC < 0, "Low1",
                                                          #ifelse(L_L2FC = 0, "Midd", 
                                                                 ifelse(L_L2FC > 0 & L_L2FC < 1, "Hig1",
                                                                        ifelse(L_L2FC > 1, "Hig2", "NA")))), from),
                              to = paste0(ifelse(R_L2FC < -1, "Low2",
                                                   ifelse(R_L2FC > -1 & R_L2FC < 0, "Low1",
                                                          #ifelse(R_L2FC = 0, "Midd", 
                                                                 ifelse(R_L2FC > 0 & R_L2FC < 1, "Hig1",
                                                                        ifelse(R_L2FC > 1, "Hig2", "NA")))), to))
sectors2 <- sort(unique(c(xx$from, xx$to)))
ordersect <- sectors2[order(match(substr(sectors2, start = 5, stop=nchar(sectors2)), substr(sectors, start = 1, stop = nchar(sectors))))]
#df.groups2 <- structure(substr(sectors2, start = 1, stop = 4), names = sectors)
Type_Cols2 <- c(Low1 = "#3399FF",Low2 = "#06038D", Hig1 = "#FF6666", Hig2 = "#E10600")

# Setup color schmee for 
y <- x %>% mutate(from = paste0(ifelse((L_L2FC & R_L2FC > 0), "Hi","Lo"), from), to = paste0(ifelse((L_L2FC & R_L2FC > 0), "Hi","Lo"), to))
sectors3 <- sort(unique(c(y$from, y$to)))
sectors3 <- sectors3[-c(16,18,41,43,44)]
ordersect3 <- sectors3[order(match(substr(sectors3, start = 3, stop=nchar(sectors3)), substr(sectors, start = 1, stop = nchar(sectors))))]
#df.groups3 <- structure(substr(sectors3, start = 1, stop =2), names = sectors)
Type_Cols3 <- c(Lo = "blue", Hi = "red")

# Figure the gaps by Grou
pdf(file = "~/Dropbox (Partners HealthCare)/SK-5A4F_SN0268732_Wesley_Nov_2022_Xenon_Human_Mouse/plots/Heatmap/HumanMicroglia_MouseAstrocytes_CircosPlots_ColAdj_redo.pdf", width = 10)
circos.clear()
circos.par(canvas.ylim = c(-1.5,1.5),track.margin = c(0.01, 0.01), track.height = (0.06))
chordDiagram(x[1:3], # Dataframe containing From and To information
             small.gap = 2, # Gap between the different genes
             big.gap = 10, # Gap between the different groups
             annotationTrack = "grid", # Only place the Grid
             annotationTrackHeight = 0.05,
             preAllocateTracks = 1,
             order = sectors, 
             directional = 1, 
             direction.type = c("diffHeight", "arrows"), 
             link.arr.type = "big.arrow",
             link.target.prop = FALSE,
             #col = j$color,
             col = structure(Type_Cols3[match(substr(sectors3, start = 1, stop = 2), labels(Type_Cols3))], names = sectors),
             #link.arr.col = "grey",
             diffHeight = -uh(1, "mm"),
             grid.col = structure(Type_Cols2[match(substr(ordersect, start = 1, stop = 4), labels(Type_Cols2))], names = sectors),
             group = df.groups)

# Add Gene Names to the Circos Plot
circos.track(track.index = 1, panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1] + 2 ,substr(CELL_META$sector.index, 3, nchar(CELL_META$sector.index)),CELL_META$sector.index, 
        facing = "clockwise", niceFacing = TRUE, adj = c(0, 0), cex = 0.6)
}, bg.border = NA)

# Highlight the different Groups
highlight.sector(x$from, track.index = 1, col = 'springgreen4', text = 'Human Microglia LIGANDS', cex = 0.6, text.col = 'white', facing = 'bending.inside', niceFacing = T)
highlight.sector(x$to, track.index = 1, col = 'magenta4', text = paste0('Astrocyte Receptors'), cex = 0.6, text.col = 'white', facing = 'bending.inside', niceFacing = T)
title("Crosstalk between Human Microglia Ligands and Mouse Astrocyte Receptors")
dev.off()
```


# 3.0 Figure 3Q
## 3.1 Generate IPA DotPlot
```{r}
##________UPDATE ENTIRE SECTION___________##
IPAgraphdir <- paste0(workdir, expdir, "plots/IPA/")
fileprefix2 <- paste0(fileprefix,"IPA_graph_pvaluesize.pdf")
##________________________________________##


# Load Data of Interst
Astrocytes6mXenvsAIR <- readxl::read_excel("~/Dropbox (Partners HealthCare)/SK-5A4G_SN0267744_Wesley_Nov_2022_Xenon/results/Madison/Astrocytes_6months/SK-5A4G_H5XFAD_6M_Astrocytes_statistics_pval01_pathways.xls", skip = 1) %>% 
  rename(c("Ingenuity.Canonical.Pathways" = "Ingenuity Canonical Pathways", "z.score" = "z-score")) %>%
  dplyr::mutate(Experiment = paste0("Astrocytes6mXenvsAIR"))

# Setup Pathways of interest
path.of.interest <- c(
"VEGF Signaling",
"STAT3 Pathway",
"JAK/STAT Signaling",
"GM-CSF Signaling",
"NRF2-mediated Oxidative Stress Response",
"IL-8 Signaling",
"NF-κB Signaling",
"White Adipose Tissue Browning Pathway",
"Apelin Adipocyte Signaling Pathway",
"RHOGDI Signaling",
"WNT/Ca+ pathway",
"Endocannabinoid Neuronal Synapse Pathway",
"Pathogen Induced Cytokine Storm Signaling Pathway")


# Generate Graph of Interest
graph <- IPApoigraph_single(data = Astrocytes6mXenvsAIR, POI = path.of.interest, directory = IPAgraphdir, fileprefix = fileprefix, suffix = "poi")

# View Graph
graph
```
